\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-04 paper/figures/example-stats.tex
words about transit and more steps for transit

\begin{floattable}
\begin{deluxetable}{cccccccc}
\caption{The scaling and convergence stats for each example. \label{tab:example-stats}}
\tablehead{\colhead{example} & \colhead{$N$} & \colhead{$J$} & \colhead{direct cost} & \colhead{\celerite\ cost} & \colhead{$D$} & \colhead{steps} & \colhead{effective samples}\\
&&& \colhead{ms} & \colhead{ms} &&&}
\startdata
1 & 200 & 1 & 2.85 & 0.26 & 3 & 80000 & 1737 \\
2 & 100 & 2 & 0.67 & 0.36 & 4 & 80000 & 1134 \\
3 & 6950 & 2 & 7351.71 & 1.46 & 4 & 176000 & 3606 \\
4 & 1440 & 8 & 828.93 & 2.74 & 9 & 640000 & 1443 \\
5 & 1000 & 2 & 82.73 & 0.94 & 13 & 640000 & 1141 \\
\enddata
\end{deluxetable}
\end{floattable}

>>>>

\begin{floattable}
\begin{deluxetable}{cccccccc}
\caption{The computational cost and convergence stats for each example. \label{tab:example-stats}}
\tablehead{& \colhead{$N$} & \colhead{$J$} & \colhead{direct\tablenotemark{a}} & \colhead{\celerite\tablenotemark{b}} & \colhead{dimension\tablenotemark{c}} & \colhead{evaluations\tablenotemark{d}} & \colhead{$N_\mathrm{eff}$\tablenotemark{e}}\\
&&& \colhead{ms} & \colhead{ms} &&&}
\startdata
1 & 200 & 1 & 2.85 & 0.26 & 3 & 80000 & 1737 \\
2 & 100 & 2 & 0.67 & 0.36 & 4 & 80000 & 1134 \\
3 & 6950 & 2 & 8119.11 & 1.47 & 4 & 176000 & 2900 \\
4 & 1440 & 8 & 828.93 & 2.74 & 9 & 640000 & 1443 \\
5 & 1000 & 2 & 88.30 & 0.96 & 13 & 1280000 & 1490 \\
\enddata

\tablenotetext{a}{The computational cost of computing the GP model using the
general Cholesky factorization routine implemented in the Intel MKL.}
\tablenotetext{b}{The computational cost of computing the GP model using
\celerite.}
\tablenotetext{c}{The total number of parameters in the model.}
\tablenotetext{d}{The total number of evaluations of the model to run MCMC to
convergence.}
\tablenotetext{e}{The effective number of independent samples estimated by
computing the integrated autocorrelation time of the chain.}
\end{deluxetable}
\end{floattable}

### 2017-06-04 paper/ms.tex
words about transit and more steps for transit

In \eq{gp-likelihood} the physical parameters of the exoplanet are called
$\bvec{\theta}$ and, in this example, the mean function $\mu_\bvec{\theta}(t)$
is a limb-darkened transit light curve \citep{Mandel:2002} and the parameters
$\bvec{\theta}$ are the orbital period $P$, the transit duration $T$, the
phase or epoch $t_0$, the impact parameter $B$, the radius of the planet in
units of the stellar radius $R_P/R_\star$, and two parameters describing
the limb-darkening profile of the star \citep{Claret:2011,Kipping:2013}.
As in \sect{rotation}, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel} and fit for the parameters of the
exoplanet $\bvec{\theta}$ and the stellar variability $\bvec{\alpha}$
simultaneously.

We take a month-long segment of the \kepler\ light curve for KIC~1430163 and
multiply it by a simulated transit model with known parameters.
The fact that we are using a simulated transit enables us to confirm that we
correctly measure the planet parameters but we do not use this knowledge to
inform our fit.
The top panel of \Figure{transit-ml} shows the data including the simulated
transit.
The bottom panel shows the same data with the maximum likelihood prediction
for the stellar variation subtracted.
To find this model, we maximized the likelihood function in \eq{gp-likelihood}
with respect to the transit parameters $\bvec{\theta}$ and the variability
parameters $\bvec{\alpha}$ simultaneously.
\Figure{transit-corner} shows the marginalized posterior constraints on the
physical properties of the planet compared to the true values.
This procedure produces estimates of the planet parameters that are consistent
with the true values using \celerite\ as an effective model for the stellar
variability.

\newpage

@@

\dfmtodo{The D in Table 6 may be confused with the D used in the Cholesky solver.}

>>>>

In \eq{gp-likelihood} the physical parameters of the exoplanet are called
$\bvec{\theta}$ and, in this example, the mean function $\mu_\bvec{\theta}(t)$
is a limb-darkened transit light curve \citep{Mandel:2002} and the parameters
$\bvec{\theta}$ are the orbital period $P_\mathrm{orb}$, the transit duration
$T$, the phase or epoch $t_0$, the impact parameter $B$, the radius of the
planet in units of the stellar radius $R_P/R_\star$, and two parameters
describing the limb-darkening profile of the star
\citep{Claret:2011,Kipping:2013}.
As in \sect{rotation}, we model the stellar variability using a GP model
with a kernel given by \eq{rot-kernel} and fit for the parameters of the
exoplanet $\bvec{\theta}$ and the stellar variability $\bvec{\alpha}$
simultaneously.
\response{The full set of parameters $\bvec{\alpha}$ and $\bvec{\theta}$ are
listed in Table~\ref{tab:example-5-params} along with their priors and the
true values for $\bvec{\theta}$.}

\response{ We take a 20~day segment of the \kepler\ light curve for
KIC~1430163 ($N = \examplevndata$) and multiply it by a simulated transit
model with the parameters listed in Table~\ref{tab:example-5-params}.
Using these data, we maximize the joint posterior defined by the likelihood in
\eq{gp-likelihood} and the priors in Table~\ref{tab:example-5-params} using
\project{L-BFGS-B} for all the parameters $\bvec{\alpha}$ and $\bvec{\theta}$
simultaneously.
The top panel of \Figure{transit-ml} shows the data including the simulated
transit as black points with the MAP model prediction over-plotted in blue.
The bottom panel of \Figure{transit-ml} shows the ``de-trended'' light curve
where the MAP model has been subtracted and the MAP mean model
$\mu_\bvec{\theta}$ has been added back in.
For comparison, the transit model is over-plotted in the bottom panel of
\Figure{transit-ml} and we see no obvious correlations in the residuals.

Sampling \examplevnwalkers~walkers from an isotropic Gaussian centered on the
MAP parameters, we run \examplevnburn~steps of burn-in and
\examplevnsteps~steps of production MCMC using \emcee.
We estimate the integrated autocorrelation time for the $\ln(P_\mathrm{orb})$
chain and find \examplevneff~effective samples across the full chain.
\Figure{transit-corner} shows the marginalized posterior constraints on the
physical properties of the planet compared to the true values.
Even though the \celeriteterm\ representation of the stellar variability is
only an effective model, the inferred distributions for the physical
properties of the planet $\bvec{\theta}$ are consistent with the true values.
This promising result suggests that \celerite\ can be used as an effective
model for transit inference for a significant gain in computational
tractability.
Even for this example with small $N$, the cost of computing the likelihood
using \celerite\ is nearly two orders of magnitude faster than the same
computation using a general direct solver.
}

@@

\dfmtodo{Do this.}
